---
title: "211012_2018_abundance_demo"
author: "Justine Albers"
date: "10/12/2021"
output: github_document
---
# Goal

This document shows how **individual bottle** bacterial abundance data from 2018 remineralization experiments were processed, QC'd and analyzed. It also provides an intro to data processing and analysis with Rstudio and R Markdown. 

#Intro to R Markdown

Create a new code chunk: Mac - Cmd + option + i, PC - Ctrl + alt + i 

Load packages that we'll need to analyze our data. 
```{r Load packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```

We can toggle on/off warnings in chunks if we don't want them in our final markdown file. 

# Import Data 

```{r}
excel_sheets("~/Documents/144l_students_2021/Input_Data/week3/144L_2018_BactAbund.xlsx")

#Mac: to make an arrow, use Option + - 

metadata <- read_excel("~/Documents/144l_students_2021/Input_Data/week3/144L_2018_BactAbund.xlsx", sheet = "Metadata")

glimpse(metadata)
#unique(metadata$Bottle)
#unique(metadata$Treatment)

data <- read_excel("~/Documents/144l_students_2021/Input_Data/week3/144L_2018_BactAbund.xlsx", sheet = "Data")
glimpse(data)

joined <-  left_join(metadata, data) #attach data to metadata
#joins right dataset to the left one by using variables that are the same across the two dataframes
glimpse(joined)

```

# Prepare Data 

We will convert the Date and Time column values from characters to dates, add columns with time elapsed for each treatment, and convert to cells/L because it will help us match up with the TOC data later. We will then subset the data for variables of interest and drop NA values. 

To do this, we are going to using **piping**. Piping is an operation that allows us to write more efficient code. The way that we'll use it here is to manipulate our data sequentially. The pipe operator "%>%", which basically says like "first do one thing to the data. THEN, do this other thing." (with the %>% operator taking the place of the word THEN in this scenario). Every call that we invoke with an additional pipe is going to take place on the variable (dataframe) that we specify at the beginning. 

```{r}
cells <- joined %>%
  mutate(Datetime = ymd_hm(Datetime), #splits apart Datetime as specified
  cells_L = Cells_ml * 1000) %>%
  group_by(Treatment, Bottle) %>%
#group our dataset so that we can calculate the time elapsed properly 
  mutate(interv = interval(first(Datetime), Datetime), 
         s = as.numeric(interv), 
         hours = s/3600, 
         days = hours/24) %>%
  ungroup() %>%
  select(Experiment:DNA_Sample, cells_L, hours, days) %>%
  drop_na(cells_L)
glimpse(cells)
```
# Plot Growth Curves

We will plot growth curve for each bottle. Need cell abundance and days data. 

First, let's set some aesthetics for our plot. 

```{r}
#assign hex colors to our different treatments
custom.colors <- c("Control" = "#377EB8", "Ash Leachate" = "#4DAF4A", "Mud Leachate" = "#E41A1C", "Glucose_Nitrate_Phosphate" = "#FF7F00")

#assign levels to control what order things appear in the legend
levels <- c("Control", "Ash Leachate", "Mud Leachate", "Glucose_Nitrate_Phosphate")

#now let's use a handy package, ggplot to visualize our data. 

cells %>%
  mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>%
  ggplot(aes(x=days, y=cells_L, group = interaction(Treatment, Bottle))) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size =1) +
  geom_point(aes(fill = factor(Treatment, levels = levels)), size = 3, color = "black", shape = 21) + 
  geom_text(aes(label = dna), size = 12, color = "#E41A1C") +
  labs(x = "Days", y = expression(paste("Cells, L"^-1)), fill = "") + 
  guides(color = "none") + 
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  #facet_grid(rows = "Treatment")
  theme_bw()
glimpse(cells)
```


# Next Steps 

We can calculate: 

- total change in cells from initial condition to the end of the experiment. 
- specific growth rate as the slope of ln(abundance) v time during exponential growth phase
- doubling time as ln(2) divided by the specific growth rate
- mean of each of these parameters across each treatment

1st, we need to determine **where** exponential growth is ocurring in each of our bottles, if it does. To do this, we 'll plot ln(abundance) vs time. 

# Identify exponential phase of growth in our remin experiments

**NOTE about logarithms in R** 

log(x) gives the natural log of x, not log base 10
log10(x) gives the log base 10
log2(x) gives log base 2

```{r}
ln_cells <- cells %>%
  group_by(Treatment, Bottle) %>%
  mutate(ln_cells = log(cells_L), 
         diff_ln_cells = ln_cells - lag(ln_cells, default = first(ln_cells)))
```

Now, let's plot our newly calculated data! 

```{r fig.height=6, fig.width=8}
ln_cells %>%
  mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>%
  ggplot(aes(x=days, y=diff_ln_cells, group = interaction(Treatment, Bottle))) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size =1) +
  geom_point(aes(fill = factor(Treatment, levels = levels)), size = 3, color = "black", shape = 21) + 
  geom_text(aes(label = dna), size = 12, color = "#E41A1C") +
  labs(x = "Days", y = expression(paste("∆ln cells, L"^-1)), fill = "") + 
  guides(color = "none") + 
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  facet_wrap("Bottle", ncol =2) +
  theme_bw()
```
Exponential growth seems to be occurring right at the beginning of the experiment between 0-1 days for the most of the bottles. 

Let's try plotting ln_cells to see if that can help us identify exponential growth in the control a little better. 

```{r fig.height=6, fig.width=8}
ln_cells %>%
  mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>%
  ggplot(aes(x=days, y=ln_cells, group = interaction(Treatment, Bottle))) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size =1) +
  geom_point(aes(fill = factor(Treatment, levels = levels)), size = 3, color = "black", shape = 21) + 
  geom_text(aes(label = dna), size = 12, color = "#E41A1C") +
  labs(x = "Days", y = expression(paste("ln cells, L"^-1)), fill = "") + 
  guides(color = "none") + 
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  facet_wrap("Bottle", ncol =2) +
  theme_bw()
```


