---
title: "test_2018_abundance"
author: "Justine Albers"
output: github_document
---
# Goal: 

# title
## title 
### title 

This document shows how **individual bottle** bacterial abundance data from 2018 remineralization experiments were processed, QC'd and analyzed. It also provides an introduction to data processing and analysis with RStudio and RMarkdown. 

# Intro to R Markdown

We will use R Markdown to make our science reproducible and easily understood by someone who is not familar with R. 

To create a new code chunk, press the green "C" button and click "R" from the dropdown menu in the toolbar above OR for Mac: type Cmd + option + i and for PC: Ctrl + alt + i.

Let's first load the packages that we'll need to analyze our data. 

```{r Load packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```

We can toggle on/off messages and warnings based on what we want to appear in our final R markdown output. 

# Import Data

We also want to set our working directory so that we can easily refer to our data files.

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Google Drive/Shared drives/EEMB144/EEMB144L_2021/R & RStudio & GitHub Info/144l_students_2021/")
```

```{r}
excel_sheets("~/")

```

